---
title: "Genetic Testing Survey Analysis"
author: "Clarissa White"
output:
html_document: default
description: null
editor_options:
chunk_output_type: console
always_allow_html: true
---
## Load libraries & data
```{r, echo=F, results='hide', warning=FALSE, message=FALSE}
setwd("~/OneDrive/Genetic_Survey")
library(MASS)
library(tidyverse) # for preprocessing
library(stringr)
library(gridExtra)
#library(zipcode) # for associating zip codes with latitude/longitude
library(caret) # create dummy variables
library(cluster) # for PAM clustering
library(e1071) # for fuzzy clustering
library(corrplot) # visualize cluster membership
library(factoextra) # visualize fuzzy clusters
library(Rtsne)
library(networkD3) # sankey diagram
library(htmltools)
# load data from Qualtrics Survey csv
data0 <- read.csv("Data/Genetic_Testing_Survey.csv", skip = 3, header = F)
# plot theme
# my_theme <- theme_bw() +
# theme(axis.text = element_text(size = 18), axis.text.x = element_text(size = 18),
# axis.title = element_text(size = 18), axis.text.y = element_text(size = 18),

# plot.title = element_text(size = 24, hjust = .5, face = "bold"))
```
## Preprocess data
```{r, echo=FALSE}
# rename columns to avoid repeats and for future convenience
colnames(data0) <- c("StartDate", "EndDate", "Status", "IPAddress", "Progress",
                     "Duration_seconds", "Finished", "RecordedDate", "ResponseId",
                    "RecipientLastName", "RecipientFirstName", "RecipientEmail",
                    "ExternalReference", "LocationLatitude", "LocationLongitude",
                    "DistributionChannel", "UserLanguage", "Q_RecaptchaScore",
                    "Consent", "Previous_testing", "Nutrition_and_Behavior",
                    "Genetic_ancestry", "Medications", "Improve_Outcome",
                    "Cannot_improve_outcome", "Active_Infection", "Employer_Test",
                    "ET_IF_who_sponsors", "ET_IF_who_conducts", "ET_IF_cost",
                    "ET_IF_returned_results", "ET_IF_sponsor_use", "ET_IF_data_access",
                    "ET_IF_data_storage_security", "ET_IF_data_ownership",
                    "ET_IF_anonymity", "ET_IF_other", "ET_other_text", "ET_comments",
                    "Insurance_Test",
                    "Ins_IF_who_sponsors", "Ins_IF_who_conducts", "Ins_IF_cost",
                    "Ins_IF_returned_results", "Ins_IF_sponsor_use", "Ins_IF_data_access",
                    "Ins_IF_data_storage_security", "Ins_IF_data_ownership",
                    "Ins_IF_anonymity", "Ins_IF_other", "Ins_other_text", "Ins_comments",
                    "Research_Study",
                    "RS_IF_who_sponsors", "RS_IF_who_conducts", "RS_IF_cost",
                    "RS_IF_returned_results", "RS_IF_sponsor_use", "RS_IF_data_access",
                    "RS_IF_data_storage_security", "RS_IF_data_ownership",
                    "RS_IF_anonymity", "RS_IF_other", "RS_other_text", "RS_comments",
                    "Employer_Test_GINA",
                    "Insurance_Test_GINA", "Employer_Test_LI",
                    "Insurance_Test_LI", "Sex", "Birth_Year",
                    "Ancestry_Asian", "Ancestry_African_Amer",
                    "Ancestry_Native_Amer", "Ancestry_Pacific_Islander",
                    "Ancestry_White", "Ancestry_Other",
                    "Ancestry_Prefer_not_to_say", "Ethnicity", "Income", "Degree",
                    "Zip_Code", "Have_Disease", "Additional_Comments")
# function to binarize word responses
binarize <- function(x) {
  out <- ifelse(trimws(x) == "", 0, 1)
  return(out)
}

data1 <- data0 %>%
  # remove unfinished and nonconsenting rows
  filter(Finished=="TRUE", Consent=="Yes") %>%
  # calculate age and list empty Sex as unknown
  mutate(Age = 2020 - Birth_Year, Sex = ifelse(Sex=="", "Did not answer", Sex)) %>%
  # binarize multiple choice responses
  mutate_at(vars(contains("IF")), binarize) %>%
  mutate_at(c(71:73,75,76), binarize) %>%
  # remove unnecessary columns
  select(-c("StartDate", "EndDate", "Status", "IPAddress", "Progress",
            "Duration_seconds", "Finished", "RecordedDate",
            "RecipientLastName", "RecipientFirstName", "RecipientEmail",
            "ExternalReference", "LocationLatitude", "LocationLongitude",
            "DistributionChannel", "UserLanguage", "Consent"))

# # create age groups
# data1$age_group <- factor(ifelse(data1$Age %in% 0:39, "< 40",
                                  # ifelse(data1$Age %in% 40:49, "40 to 49",
                                          # ifelse(data1$Age %in% 50:59, "50 to 59", "60+"))),
                            # levels = c("< 40", "40 to 49", "50 to 59", "60+"))

# column of ones for getting group counts
data1$counts = as.numeric(1)
# interested in at least 1 test, no tests, or all tests
data1$test_interest <- ifelse((data1$Nutrition_and_Behavior == "Yes")
                              & (data1$Genetic_ancestry == "Yes")
                              & (data1$Medications == "Yes")
                              & (data1$Improve_Outcome == "Yes")
                              & (data1$Cannot_improve_outcome == "Yes")
                              & (data1$Active_Infection == "Yes"), "All",
                              ifelse((data1$Nutrition_and_Behavior == "Yes")
                                     | (data1$Genetic_ancestry == "Yes")
                                     | (data1$Medications == "Yes")
                                     | (data1$Improve_Outcome == "Yes")
                                     | (data1$Cannot_improve_outcome == "Yes")
                                     | (data1$Active_Infection == "Yes"), 
                                     "At Least One", "None"))
head(data1[,1:3])
```

## Explore demographics of survey participants
```{r, fig.width=9, fig.height=5, echo=FALSE, warning=FALSE, message=FALSE}
## general demographics
# age histogram
ggplot(data1, aes(x = as.numeric(as.character(Age)))) +
  geom_histogram(binwidth = 5, color = "black", fill = "#003C71") +
  ggtitle("Age Distribution") + xlab("Age") + ylab('Count')

# sex bar plot
ggplot(data1, aes(x = Sex, y = counts)) +
  geom_bar(stat = "identity", fill = "#003C71") +
  ggtitle("Sex Distribution") + ylab('Count')
# ancestry
ggplot(as.data.frame(t(data1 %>% select(Ancestry_Asian:Ancestry_Other) %>%
                         summarize_all(list( ~ sum(.))))) %>%
         mutate(Ancestry = substr(rownames(.), 10, nchar(rownames(.)))),
       aes(x = fct_reorder(Ancestry, V1, .desc = T), y = V1)) +
  geom_bar(stat = "identity", fill = "#003C71") +
  ggtitle("Ancestry Distribution") +
  xlab("Ancestry") + ylab("Count") + 
  theme(axis.text.x = element_text(size = 18, angle = 20, vjust=.6))
# ethnicity bar plot (should I make this a proportion?) #########
ggplot(data1[data1$Ethnicity!="",], aes(x = Ethnicity, y = counts)) +
  geom_bar(stat = "identity", fill = "#003C71") +
  ggtitle("Ethnicity Distribution") +
  theme(axis.text.x = element_text(size = 18, angle = 20, vjust=.6))
# income (Sort by income not number of people with that income) #########
ggplot(data1[data1$Income!="",] %>% group_by(Income) %>% summarize(count =
  sum(counts)),
  aes(x = fct_reorder(Income, count, .desc = T), y = count)) +
  geom_bar(stat = "identity", fill = "#003C71") +
  ggtitle("Income Distribution") +
  xlab("Income") + ylab("Count") +
  theme(axis.text.x = element_text(size = 12, angle = 20, vjust=.6))

# highest degree
ggplot(data1[data1$Degree!="",] %>% group_by(Degree) %>% 
         summarize(count = sum(counts)),
       aes(x = fct_reorder(Degree, count, .desc = T), y = count)) +
  geom_bar(stat = "identity", fill="#003C71") + xlab("Degree") + ylab("Count") +
  ggtitle("Highest Degree Earned") +
  theme(axis.text.x = element_text(size = 18, angle = 20, vjust=.6))

# zip code #########
# data(zipcode)
# ggplot(fm, aes(longitude, latitude)) +
# geom_polygon(data = us, aes(x = long, y = lat, group = group), color='gray', fill=NA,alpha=.35)
+
# geom_point(aes(color = count),size = 0.15,alpha = 0.25) +
# xlim(-125, -65) + ylim(20, 50)
```

## Plot and test using Fisher's Test whether Sex is associated with interest in different genetic
tests
```{r, fig.width=8, fig.height=5, echo=FALSE, warning=FALSE, message=FALSE}
# function to perform chi-squared test on sex and test type interest

## (could be extended to other demographics)
bar_plot_associations <- function(test_type, test_chr, dem_var = "Sex", dem_chr = "Sex") {
  # data
  # remove non-female/male categories
  if (dem_chr == "Sex") {
  plot_data <- select(data1[which(data1$Sex %in% c("Female", "Male")),],
  counts, all_of(dem_var), all_of(test_type))
  } else {
  plot_data <- select(data1, counts, all_of(dem_var), all_of(test_type))
  }
  colnames(plot_data) <- c("counts", "dem_var", "test_type")
  
  # chi-squared test
  my_tbl <- table(plot_data$dem_var, plot_data$test_type)
  pval <- fisher.test(my_tbl)$p.value
  
  # plot
  ggplot(plot_data %>% group_by(dem_var, test_type) %>% count(counts) %>%
            ungroup() %>% group_by(dem_var) %>%
            filter(test_type!="") %>%
            mutate(freq = round(n/sum(n), 2), tot = cumsum(n),
                   pos = ifelse(freq < .5, 1 - (0.5 * freq), freq - (.45 * freq))),
         aes(x = dem_var, y = freq, fill = test_type)) +
    # plot type and axis labels
    geom_bar(stat='identity') + xlab(dem_chr) + ylab("Proportion") + # my_theme +
    # add frequency labels to the bars
    geom_text(aes(x = dem_var, y = pos, label = freq), size = 8) +
    guides(fill=guide_legend(title = "Interested")) +
    # add count of participants in each bar
    # geom_text(aes(y = 1, label = c("N = 27","","N = 28","","N = 22","", "N = 33", ""),
    # vjust = -.5), size = 7) +
    ggtitle(paste("Interested in Genetic Test About\n", test_chr,
            paste(" (p-value = ", round(pval, digits = 4), ")", sep = ""), sep = ""))
}

# Sex
bar_plot_associations("Nutrition_and_Behavior", "Nutrition and Behavior")
bar_plot_associations("Genetic_ancestry", "Genetic Ancestry")
bar_plot_associations("Medications", "Drug Response")
bar_plot_associations("Improve_Outcome", "a Treatable Disease")

bar_plot_associations("Cannot_improve_outcome", "an Untreatable Disease")
bar_plot_associations("Active_Infection", "an Active Infection")

# Interested in genetic tests at all
my_tbl <- table(data1$Sex, data1$test_interest)
pval <- fisher.test(my_tbl)$p.value
ggplot(data1[which(data1$Sex %in% c("Female", "Male")),] %>%
         group_by(Sex, test_interest) %>%
         summarise(n = n()) %>%
         mutate(freq = round(n/sum(n), 2)),
       aes(x = Sex, y = freq, fill = test_interest)) +
  geom_bar(stat = 'identity') + xlab("Sex") + ylab("Proportion") +
  guides(fill = guide_legend(title = "")) +
  ggtitle(paste("Interested in Genetic Tests At All\n",
  paste("(p-value = ", round(pval, digits = 4), ")", sep = "")))
```

## Plot and test whether Age is associated with interest in different genetic tests
```{r, fig.width=6.5, fig.height=5, echo=FALSE, warning=FALSE, message=FALSE}
# function to perform ttest on age and test type interest
box_plot_associations <- function(test_type, test_chr, dem_var = "Age", dem_chr = "Age") {
  # data
  plot_data <- select(data1, dem_var, test_type)
  colnames(plot_data) <- c("dem_var", "test_type")
  # t-test
  pval <- t.test(plot_data %>% filter(test_type=="Yes") %>% select(dem_var),
                 plot_data %>% filter(test_type=="No") %>% select(dem_var), 
                 var.equal = FALSE)$p.value
  # plot
  ggplot(plot_data %>% 
           filter(test_type %in% c("Yes", "No")) %>%
           group_by(test_type) %>%
           mutate(n = n()),
         aes(x = test_type, y = dem_var, fill = test_type)) +
    # plot type and axis labels
    geom_boxplot() + xlab("Interested") + ylab(dem_chr) + # my_theme +
    # add count of participants in each group
    geom_text(aes(x = test_type, y = 95, label = paste("N =", n)), size = 5) +
    theme(legend.position = "none") +
    ggtitle(paste("Interested in Genetic Test About\n", test_chr,
    paste(" (p-value = ", round(pval, digits = 4), ")", sep = ""), sep = ""))

}

# Age
box_plot_associations("Nutrition_and_Behavior", "Nutrition & Behavior")
box_plot_associations("Genetic_ancestry", "Genetic Ancestry") # * significant
box_plot_associations("Medications", "Drug Response")
box_plot_associations("Improve_Outcome", "a Treatable Disease")
box_plot_associations("Cannot_improve_outcome", "an Untreatable Disease") # * significant
box_plot_associations("Active_Infection", "an Active Infection") # * significant

# Plot how many people were interested in at least one, none, or all tests
# (all being a subset of at least one)
# Age continuous
ggplot(data1, aes(x = test_interest, y = Age, fill = test_interest)) +
  geom_boxplot() +
  xlab("Interest") + ylab("Age") +
  theme(legend.position = "None") +
  ggtitle("Interested in Genetic Tests At All")
knitr::knit_exit()
```

## Plotting level of interest before and after learning about GINA or life insurance policy
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# keep only necessary variables
data2 <- data1 %>% select(counts, Employer_Test, Employer_Test_GINA, Employer_Test_LI,
Insurance_Test, Insurance_Test_GINA, Insurance_Test_LI)

# Sankey for changing answer
plot_sankey <- function(test_provider, change, char = "GINA") {
# get direction of answer changes 3x3 table
ans_change <- ftable(data2[[test_provider]], data2[[change]])
ans_change_2 <- ans_change[(nrow(ans_change)-2):nrow(ans_change),
                           (ncol(ans_change)-2):ncol(ans_change)]
# Possible before and after answers
nodes = data.frame("Interest" = c("Unsure", "No", "Yes", "Unsure", "No", "Yes"))
# Number of participants going from each answer to the next
# (for example: 0, 3, 68 means 68 participants said no before and after)

links = as.data.frame(matrix(c(
0, 3, ans_change_2[1,1], # unsure to unsure
0, 4, ans_change_2[1,2], # unsure to no
0, 5, ans_change_2[1,3], # unsure to yes
1, 4, ans_change_2[2,2], # no to no
1, 3, ans_change_2[2,1], # no to unsure
1, 5, ans_change_2[2,3], # no to yes
2, 5, ans_change_2[3,3], # yes to yes
2, 3, ans_change_2[3,1], # yes to unsure
2, 4, ans_change_2[3,2]), # yes to no
byrow = TRUE, ncol = 3))
names(links) = c("source", "target", "value")
# Add a 'group' column to each connection:
links$group <- as.factor(c("unsure","unsure", "unsure",
                           "no","no", "no",
                           "yes", "yes", "yes"))
# Give a color for each group:
my_color <- 'd3.scaleOrdinal() .domain(["unsure", "no", "yes"]) .range(["#FCAE00", "#009104",
"#003C71"])'
# Perform McNemar's Test to get a p-value
pval <- mcnemar.test(ans_change_2, correct = F)$p.value
# Plot
# change this so it plots in plot section
p <- browsable(
  tagList(
  tags$h1(paste("Change in Interest in", test_provider, "After Learning About", char,
  paste("(p-value = ", format(pval, digits = 2, scientific = T), ")", sep = ""))),
  sankeyNetwork(Links = links, Nodes = nodes, Source = "source",
  Target = "target", Value = "value", NodeID = "Interest",
  fontSize= 58, nodeWidth = 100, colourScale = my_color,
  LinkGroup = 'group', iterations = 0)
  )
  )
  #return(list(ans_change, mcnemar.test(ans_change, correct=F), p))
  #return(list(ans_change_2, p))
  return(p)
}

# Employer Test
## GINA
plot_sankey("Employer_Test", "Employer_Test_GINA")
## Life Insurance
plot_sankey("Employer_Test", "Employer_Test_LI", "Life Insurance")
# Medical Insurance
## GINA
plot_sankey("Insurance_Test", "Insurance_Test_GINA")
## Life Insurance
plot_sankey("Insurance_Test", "Insurance_Test_LI", "Life Insurance")

## Still to do: look at which qualitative tags most influenced changes in answer
```

## Most important factors
```{r, echo=FALSE}
# employer test
ET_IF <- as.data.frame(t(data1 %>%
                           select(Employer_Test, ET_IF_who_sponsors:ET_IF_other) %>%
                           group_by(Employer_Test) %>%
                           summarise_all(sum)))
et_total <- data1 %>% group_by(Employer_Test) %>% count()
# other
data1[which(data1$ET_IF_other==1),]$ET_other_text

# insurance test
Ins_IF <- as.data.frame(t(data1 %>%
                            select(Insurance_Test, Ins_IF_who_sponsors:Ins_IF_other) %>%
                            group_by(Insurance_Test) %>%
                            summarise_all(sum)))
ins_total <- data1 %>% group_by(Insurance_Test) %>% count()
# other
data1[which(data1$Ins_IF_other==1),]$Ins_other_text

# research study
RS_IF <- as.data.frame(t(data1 %>% 
                           select(Research_Study, RS_IF_who_sponsors:RS_IF_other) %>%
                           group_by(Research_Study) %>%
                           summarise_all(sum)))
rs_total <- data1 %>% group_by(Research_Study) %>% count()
# other
data1[which(data1$RS_IF_other==1),]$RS_other_text
```
